---
title: "QTM 151"
subtitle: "Week 11 -- maps <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: ""
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      self-contained: false
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "https://cloud.r-project.org/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>

# Recap

.font130[

We learned:
 + `qplot`: quick way to make ggplot graphs.
 + `ggplotly` and `plot_ly`: create nice plotly graphs.
 + `dplyr` methods: data wrangling
 + `dplyr *_join` methods: joining data
 + `tidyr` methods: reshape datasets
 + `forcats` methods: working with categorical variables
 + `lubridate` methods: processing dates and times

**Great job!!** Do you have any questions about any of these contents?

Today we are going to talk about how to make maps in R.

Our GitHub page is: https://github.com/umbertomig/qtm151
]

---

class: inverse, center, middle

# Getting Started

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Getting Started: loading packages

.font130[
```{r}
# Loading tidyverse
library(tidyverse)
library(maps)
library(ggmap)
library(nycflights13)
```
]

---


# Getting Started: loading data

.font130[
```{r}
GAdat<-read.csv("https://raw.githubusercontent.com/umbertomig/qtm151/main/datasets/GAdat.csv")
GAdat$County <- tolower(GAdat$County)
head(GAdat, 2)
```
]

---

class: inverse, center, middle

# ggmap

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# ggmap package

.font130[

The package makes it easy to retrieve raster map tiles from popular online mapping services like Stamen Maps and Google Maps

It is easy to plot the info using the ggplot2 framework.

Suppose we want to plot the average arrival delay of NY flights.

```
data1<-flights %>%
  drop_na() %>%
  group_by(dest) %>%
  summarise(average=mean(arr_delay)) %>%
  left_join(airports, by=c("dest"="faa"))
head(data1, 2)
```
]

---

# ggmap package

.font130[

```
us <- c(left = -125, bottom = 25.75, right = -67, top = 49)
map <- get_stamenmap(us, zoom = 5, maptype = "toner-lite")
ggmap(map)  # create a layer US map

ggmap(map) + geom_point(data=data1, 
    aes(x=lon, y=lat, color=average, size=average), na.rm = T) +
  scale_color_gradient(low = "green", high="darkblue")
```

**Your turn**: do the same graph, but with the variable `distance`, instead of `arr_delay`. You should learn smt trivial!
]

---

# ggmap package

.font130[

Let's now plot the number of flights from each destination.

```
data2<-flights %>%
  drop_na() %>%
  group_by(dest) %>%
  summarise(sum=n()) %>%
  left_join(airports, by=c("dest"="faa")) 
head(data2, 2)
```
]

---

# ggmap package

.font130[

```
us <- c(left = -125, bottom = 25.75, right = -67, top = 49)
map <- get_stamenmap(us, zoom = 5, maptype = "toner-lite")
ggmap(map)

ggmap(map) + geom_point(data=data2,
        aes(x=lon, y=lat, color=sum, size=sum), na.rm = T) +
  scale_color_gradient(low = "blue", high="red")
```

**Your turn**: Plot the number of flights for the *Delta* (DL) carrier only.

]

---

class: inverse, center, middle

# maps package

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# maps package

.font130[

It allows us to turn data from the maps into a data frame suitable for plotting with ggplot.

The structure of the data needed:

* `long`: longitude.
* `lat`:  latitude.
* `order`: shows in which order ggplot should “connect the dots”
* `region` and `subregion`: tell what region or subregion a set of points surrounds.
* `group`: ggplot2’s functions can take a group argument, which controls whether adjacent points should be connected by lines. 
  + If they are in the same group, then they get connected, but if they are in different groups then they don't.

]

---

# maps package

Plot the USA map using *geom_polygon()*.

*geom_polygon()* drawn lines between points and “closes them up” (i.e. draws a line from the last point back to the first point). 

You have to map the group aesthetic to the group column.

.font130[

```
# map_data function in maps package
states<-map_data("state")  
head(states)
qplot(long, lat, data=states)
qplot(long, lat, data=states, geom="path")
```

* And they look ugly...
]
---

# maps package

To make them look better, we need to group!

.font130[

```
qplot(long, lat, data=states, geom="path", group=group)

qplot(long, lat, data=states, geom="polygon", group=group)

# color for boarder lines
ggplot(states)+
  geom_polygon(aes(x=long, y=lat, group=group), color="red") 

# fill for inside color
qplot(long, lat, data=states, geom="polygon", 
  group=group, fill=long, color="red")
```

* And they look much better.
]
---

# maps package

.font130[

We can also turn off the color legend:

```
states <- map_data("state")
ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + 
  coord_fixed(1.3) +
  guides(fill=FALSE)
```
]
---

# maps package

.font130[

We can also plot a subregion:

```
west_coast <- subset(states, region %in% c("california", "oregon", "washington"))

ggplot(data = west_coast) + 
  geom_polygon(aes(x = long, y = lat), fill = "palegreen", 
    color = "black")
```
]
---

# maps package

.font130[

But we can do better:

* group() 

* coord_fixed(): it fixes the relationship between one unit in the y direction and one unit in the x direction. 
 + Every y unit was 1.3 times longer than an x unit, the plot came out looking good.

```
ggplot(data = west_coast) + 
  geom_polygon(aes(x = long, y = lat, group = group), 
    fill = "palegreen", color = "black") + 
  coord_fixed(1.3)
```

**Your Turn**: Do the same plot for a different set of states of your choice.
]
---


# maps package

.font130[

And to plot only one state, we need to filter this state out of the dataset:

```
states <- map_data("state")

ga_df <- states %>%
  filter(region == "georgia")

ggplot(data = ga_df) + 
  geom_polygon(aes(x = long, y = lat), fill = "palegreen", color = "black")
    
ggplot(data = ga_df) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill = "palegreen", color = "black") + 
  coord_quickmap()
```

**Your Turn**: Do the same plot for a different US state.
]
---

# maps package

.font130[

Using theme_void(): you can make a plot with no background.

```
ga_base <- ggplot(data = ga_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_quickmap() + 
  geom_polygon(color = "black", fill = "gray")

ga_base + theme_void()
```
]
---


# maps package

.font130[

The package maps also have subregions and subdivisions within a given country.

This is handful to plot counties in Georgia:

```
county_df <- map_data("county") %>% filter(region == "georgia")
state_df <- map_data("state") %>% filter(region == "georgia")

ga_base + theme_void() + 
  geom_polygon(data = county_df, fill = NA, color = "white") +
  geom_polygon(color = "black", fill = NA) 
```
]
---


# maps package

.font130[

And in the GAdat we have the population by county. We can plot this in the Georgia map:

```
county_df$subregion <- replace(county_df$subregion, county_df$subregion=="de kalb", "dekalb")

mapdat <- left_join(GAdat,county_df, by = c("County"="subregion"))

p <- ggplot(mapdat, aes(long, lat, group = group)) +
    geom_polygon(aes(fill = Population), color="yellow")  +
    scale_fill_gradient(low="blue", high="red")+
    geom_polygon(data = state_df, colour = "black", fill = NA) + 
    theme_void() +
    coord_fixed(1.2) 
p
```
]
---

# maps package

.font130[

And for an alternative map:

```
ggplot(mapdat, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = Population, color="yellow"),
    colour = alpha("red", 1/2))  +
  geom_polygon(data = state_df, colour = "black", fill = NA) + 
  theme_void() + coord_fixed(1.2) +
  scale_fill_gradientn(colours = rev(rainbow(7)),
   breaks = c(2, 4, 10, 100, 1000, 10000), trans = "log10")
```
]
---


class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>

---

class: inverse, center, middle

# Have a great weekend!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>