---
title: "QTM 151"
subtitle: "Week 10 -- plotly (cont'd) <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: ""
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "https://cloud.r-project.org/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>

# Recap

.font130[

We've learned:
 + Programming methods
 + `dplyr *_join` methods: joining data
 + `tidyr` methods: reshape datasets
 + `forcats` methods: working with categorical variables
 + `lubridate` methods: to work with dates
 + `qplot` methods: to create quick plots
 + `ggplotly -- plotly` methods: create quick interactive plots.
 
Do you have any questions about `ggplotly`?

The quiz is posted. Good luck!

Our GitHub page is: https://github.com/umbertomig/qtm151
]

---

# Today's Agenda

.font150[

* `plot_ly` today:
  - Scatter-plots
  - Line-plots
  - Bar-plots

]

---

class: inverse, center, middle

# Getting Started

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Getting Started: loading packages

.font130[
```{r}
# Loading tidyverse
library(tidyverse)

# Loading plotly
library(plotly)

# Load gapminder
library(gapminder)
```
]

---

# Loading datasets

.font130[
```{r, tidy=FALSE}
# Loading tips dataset
tips <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm151/main/datasets/tips.csv') # or your locale folder
head(tips, 2)

# Loading PErisk dataset
PErisk <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm151/main/datasets/PErisk.csv') # or your locale folder
head(PErisk, 2)
```
]

---

# Loading datasets

```{r, tidy=FALSE}
# Gapminder
head(gapminder, 2)
```

---
class: inverse, center, middle

# plot_ly: scatter-plots

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# `plot_ly` Scatter-plots

.font130[

* It is very simple to build a plotly scatter-plot.

* The most basic syntax is the following:
```
  plot_ly(data = data_set_here, 
    x = ~x_axis_here, 
    y = ~y_axis_here,
    type = 'scatter')
```

- Let's try?!

]

---

# plotly

```
head(PErisk)
plot_ly(data = PErisk, 
  x = ~barb2, 
  y = ~gdpw2)
```

---

# plot_ly Scatter-plots styling

.font130[
* We can change the style of our plot_ly. For example: we can change the dots and add text to markers:
]

```
plot_ly(data = PErisk, x = ~barb2, y = ~gdpw2,
        text = ~paste("Country: ", country), # Hover text
        type = 'scatter',
        marker = list(size = 10, # Markers characteristics
                      color = 'red',
                      line = list(color = 'black',
                                  width = 2))) %>%
  layout(title = 'Informal Markets and GDP', # Layout
         yaxis = list(zeroline = FALSE),
         xaxis = list(zeroline = FALSE))
```

---

# `plot_ly` Scatter-plots qualitative coloring

.font150[
* We can also color our scatter-plot by a qualitative variable:
]

```
PErisk$courts2 <- factor(PErisk$courts, labels = c('Not Independent','Independent'))
plot_ly(data = PErisk, x = ~barb2, y = ~gdpw2,
        text = ~paste("Country: ", country), # Hover text
        color = ~courts2, colors = c('blue', 'red'),
        type = 'scatter', mode = 'markers',
        marker = list(size = 10)) %>%
  layout(title = 'Informal Markets and GDP', # Layout
         yaxis = list(zeroline = FALSE),
         xaxis = list(zeroline = FALSE))
```
---

# `plot_ly` 3D-Scatter-plots

.font130[
* It is also easy to build 3D scatterplots.

* Here we plot infant mortality and GDP in the US by year.

* Warning: the plot is very awkward.
]

```
gmd <- filter(gapminder, country == 'United States')
plot_ly(data = gmd, x = ~year, y = ~gdpPercap, z = ~lifeExp,
        color = ~year,
        type = 'scatter3d', mode = 'markers',
        marker = list(size = 10)) %>%
  layout(title = 'U.S. GDP and Life Expectancy by Year', # Layout
         yaxis = list(zeroline = FALSE),
         xaxis = list(zeroline = FALSE))
```
---

# `plot_ly` Scatter-plots

.font150[
* **Your turn:** Create a `plot_ly` with `totbill` and `tip`. Then, change the title, remove the x and y zerolines, and display the weekday (`day`) in the text. Color by `time`.
]

---

class: inverse, center, middle

# plot_ly: line-plots

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# `plot_ly` Line Plots

.font130[

* Line plots are very useful for time series, such as stock prices.

* The most basic syntax is the following:
```
  plot_ly(data = data_set_here, 
    x = ~x_axis_here, 
    y = ~y_axis_here, 
    type = 'scatter', 
    mode = 'lines')
```

- Let's try?!

]

---

# plot_ly: line-plots

```
gmd <- filter(gapminder, country == 'United States')
plot_ly(data = gmd, x = ~year, y = ~gdpPercap,
    type = 'scatter', 
    mode = 'lines') %>%
  layout(title = 'U.S. GDP per capita over time', # Layout
         yaxis = list(title = 'GDP Per Capita'),
         xaxis = list(title = 'Year'))
```

---

# plot_ly: line-plots

.font130[
* We can add markers to the lines:
]

```
plot_ly(data = gmd, x = ~year, y = ~gdpPercap,
    type = 'scatter', 
    mode = 'lines+markers') %>%
  layout(title = 'U.S. GDP per capita over time', # Layout
         yaxis = list(title = 'GDP Per Capita'),
         xaxis = list(title = 'Year'))
```

---

# plot_ly: line-plots with multiple lines

.font130[
* We can also add different countries:
]

```
gmd <- filter(gapminder, country %in% c('United States', 'Canada'))
plot_ly(data = gmd, x = ~year, y = ~gdpPercap,
  type = 'scatter', linetype = ~country,
  mode = 'lines+markers') %>%
  layout(title = 'U.S. and Canada GDP per capita over time',
         yaxis = list(title = 'GDP Per Capita'),
         xaxis = list(title = 'Year'))
```

---

# plot_ly: line-plots with multiple lines

.font130[
* We can also plot all countries in the same chart.
]

```
plot_ly(data = gapminder, x = ~year, y = ~log(gdpPercap)) %>%
  add_lines(color =~country) %>%
  hide_legend()
```
---

# plot_ly: line-plots with multiple lines

.font130[
* And change the color of a desired country.
]

```
gapminder %>%
  group_by(country) %>%
  plot_ly(x = ~year, y = ~log(gdpPercap)) %>%
  add_lines(text = 'All Countries', alpha=I(0.2)) %>%
  filter(country=="United States") %>%
  add_lines(name="United States", color=I("blue")) %>%
  add_lines(data = filter(gapminder, country=="Canada"),
    name="Canada", color=I("red"))
```
---

# plot_ly: line-plots

.font150[
* **Your turn:** In the `gapminder` dataset, select three countries of your choice and plot them.
]
---

class: inverse, center, middle

# plot_ly: bar-plots

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# `plot_ly` Bar-plots

.font130[

* Bar plots are great to visualize qualitative data.

* The most basic syntax is the following:
```
  plot_ly(data = data_set_here, 
    x = ~x_axis_here, 
    y = ~y_axis_here, 
    type = 'bar')
```

- Let's try?!

]

---

# plot_ly: bar-plots

```
tbl <- table(PErisk$prsexp2); tbl <- as.data.frame(tbl)
plot_ly(data = tbl, x = ~Var1, y = ~Freq,
  type = 'bar') %>%
  layout(title = 'Expropriation Risk in 1992', # Layout
         xaxis = list(title = 'Expropriation Risk (high to low)'))
```

---

# plot_ly: bar-plots

.font130[
* We can add multiple bars to the barplot:
]

```
tbl <- table(PErisk$prsexp2); tbl <- as.data.frame(tbl)
tbl2 <- table(PErisk$prscorr2); tbl2 <- as.data.frame(tbl2)
dat <- data.frame(Risk = tbl$Var1, exprop = tbl$Freq, corrup = tbl2$Freq)
plot_ly(data = dat, x = ~Risk, y = ~exprop,
  type = 'bar', name = 'Expropriation Risk') %>%
  add_trace(y = ~corrup, name = 'Corruption Risk') %>%
  layout(title = 'Investment Risks in 1992', # Layout
         yaxis = list(title = 'Frequencies'),
         xaxis = list(title = 'Risks'),
         barmode = 'group')
```

---

# plot_ly: bar-plots

.font150[
* **Your turn:** In the `tips` dataset, create a barplot with gender and days that the person goes to a pub.
]
---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>

---

class: inverse, center, middle

# Have a great weekend!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>