---
title: "QTM 151"
subtitle: "Lab 05 -- lubridate <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: "Summer"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      self-contained: false
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "https://cloud.r-project.org/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>

# Recap

.font130[

We learned:
 + `qplot`: quick way to make ggplot graphs.
 + `ggplotly` and `plot_ly`: create nice plotly graphs.
 + `dplyr *_join` methods: joining data
 + `forcats` methods: working with categorical variables

**Great job!!**

Do you have any questions about any of these contents?

Today we are going to talk about **lubridate** (package for dates and time)

]

---

# This class

.font130[

We have a **quiz** due by the next class.

We have the next batch of DataCamp Assignments due soon.

Our GitHub page is: https://github.com/umbertomig/qtm151
]

---

class: inverse, center, middle

# Getting Started

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Getting Started: loading packages

.font130[
```{r}
# Loading tidyverse
library(tidyverse)
library(lubridate)
```
]

---

class: inverse, center, middle

# lubridate

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# lubridate

.font130[
Provide tools to work with dates. The methods we will use in here are:

* `ymd()`, `ydm()`, `mdy()`, `myd()`, `dmy()`, `dym()`: read text into data, provided in this order.
* `parse_date_time()`: parse other date and time objects.
* `make_date()` and `make_datetime()`: parse dates and times.
* `round_dates()` and others: round dates and times.
* `difftime()`: time/dates differences.
* `now()` and `today()`: current dates and times.
* `seconds()`, `dseconds()` and others: create time spans and durations.
* `%--%` and date intervals: create and check date intervals.

]

---

class: inverse, center, middle

# Dates and Strings

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Dates and Strings

.font130[

Without declare an object as a date, it is just a string.

Check that in R (for ISO 8601):

```
dt <- '2021-02-22'
class(dt)
str(dt)

dt <- as.Date(dt)
class(dt)
str(dt)
```
]


---

class: inverse, center, middle

# ymd() and related functions

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# ymd() and related functions

.font130[

These functions are great to parse dates:

```
x <- "2010 September 20th"
ymd(x)

y <- "02.01.2010"
dmy(y)

z <- "Sep, 12th 2010 14:00"
mdy_hm(z)

a <- "2016-07-08 12:34:56"
ymd_hms(a)
```
]

---

# ymd() and related functions

.font130[

**Your turn:** Turn into date:

1. "2010-01-22"
2. "12-02-1234 12:30:25"
3. "May 18, 1933"
4. "1st June 2012"

]

---

class: inverse, center, middle

# parse_date_time

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# parse_date_time

.font130[

`parse_date_time()` takes two arguments: First, a vector with dates. Second, a vector with orders to parse.

E. g.: `parse_date_time(date = "some_date", order = "some_order")`

Some patterns:

- `dmy`: day, month, year (4-digit)
- `mdY`: month, day, year (2-digit)
- `ymdHMS`: year (4-digit), month, day, hours, minutes, seconds

For more: `help(parse_date_time)`
]

---

# parse_date_time

.font130[

Examples:

```
parse_date_time(c("11-01-2016", "2016 Jan 11th"),
  order = c("dmy", "ymd"))
parse_date_time("11-01-2016", order = "dmy")
```

**Your turn:** Parse:

1. `x <- c("April 13, 2003", "17 April 2005")`
2. `y <- "January 10, 2020 at 23:30:35"`

]

---

class: inverse, center, middle

# make_date() and make_datetime()

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# make_date() and make_datetime()

.font130[

Both functions work to build date and time from pieces. The syntax for both functions is straightforward:

```
make_date(year = 2012, month = 3, day = 27)

make_datetime(year = 1234, month = 5, day = 12, 
  hour = 1, min = 23, sec = 45)
```

**Your turn:** make dates from:
- yrs <- c(1234, 1222, 2020)
- mos <- c(4, 2, 12)
- dss <- c(12, 12, 20)

]

---

class: inverse, center, middle

# Extract parts functions

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Extract parts functions

.font130[

- year(): Year with century
- month(): Month
- day(): Day of month
- hour(): Hour
- min(): Minute
- second(): Second
- wday(): Weekday
- yday(): Day of year
- tz(): Timezone
- quarter(): Quarter
- semester(): Semester

]

---

# Extract parts functions

.font150[

**Your turn:** Parse the dates and try these functions with:

1. "2010-01-22"
2. "12-02-1234 12:30:25"

]

---

class: inverse, center, middle

# round_dates() and others

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# round_dates() and others

.font130[

To round dates we can use:

- `round_dates()`: round to the nearest precision.
- `ceiling_date()`: round up the nearest precision.
- `floor_dates()`: round down the nearest precision.

Precision units:

- "second" , "minute" , "hour" , "day" , "week" , "month",
"bimonth" , "quarter" , "halfyear" , "year", or multiples (e.g. "10 minutes")

]

---

# round_dates() and others

.font130[

Examples:

```
d <- ymd_hms("1234-04-03 07:13:28 UTC")
floor_date(d, unit = 'day')

# Round to nearest 5 minutes
round_date(d, unit = '5 minutes')

# Round up to week 
ceiling_date(d, unit = 'week')

# Fun
ceiling_date(d, unit = 'hour') - floor_date(d, unit = 'hour')
```

**Your turn:** Round `d` in the three different ways: `d <- ymd_hms("1981-08-18 18:08:28 UTC")`

]

---

class: inverse, center, middle

# difftime

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# difftime

.font130[

We can find difference between two dates using difftime.

```
difftime(today(), mdy("May 20, 2020"), units = 'days')
difftime(now(), mdy_hms("May 20, 2020 22:12:22"), units = 'secs')
```

**Your turn:** Find how many seconds since your birthday.

]

---

class: inverse, center, middle

# seconds(), dseconds() and other duration functions

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# seconds(), dseconds() and other duration functions

.font130[

We can use duration to change dates:

```
today() - years(1)
today() + month(1)
days(2)
ddays(2)
now() + seconds(20)
2*days()
```

**Your turn:** Find how many seconds since your birthday.

]

---

class: inverse, center, middle

# Intervals

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Intervals

.font130[

We can also compute intervals:

```
x <- dmy("2 January 1998") %--% dmy("30 March 2018")
y <- dmy("2 January 2017") %--% dmy("30 March 2020")
int_start(x)
int_end(x)
int_length(x)
as.period(x)
dmy("5 January 1998") %within% x
dmy("5 January 1998") %within% y
int_overlaps(x, y)
```

**Your turn:** Compute one interval between your birthday and today. Check it. Does it contains the date `January 20 2000`?

]

---


class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>

---

class: inverse, center, middle

# See you next class!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>