---
title: "QTM 151"
subtitle: "Week 5 -- dplyr <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: ""
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "https://cloud.r-project.org/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>

# Recap

.font130[

* We learned:
 + `qplot`: quick way to make ggplot graphs.
 + `ggplotly`: transform ggplot objects into nice plotly viz.
 + `plot_ly`: create nice plotly graphs.

* Do you have any questions about any of these functions?

* Quiz is posted!

* Our GitHub page is: https://github.com/umbertomig/qtm151
]

---

# Today's Agenda

.font130[

`dplyr` package for data manipulation: 
* Observation/row:
    + filter
    + arrange
* Variable/column:
    + select
    + rename
    + mutate
* Get summaries:
    + summarise
    + group_by: changes the scope of each function from operating on the entire dataset to operating on it group-by-group

]

---

class: inverse, center, middle

# Getting Started

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Getting Started: loading packages

.font130[
```{r}
# Loading tidyverse
library(tidyverse)
```
]

---

# Loading datasets

.font130[
```{r, tidy=FALSE}
# Loading tips dataset
tips <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm151/main/datasets/tips.csv') # or your locale folder
head(tips, 2)

# Loading PErisk dataset
PErisk <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm151/main/datasets/PErisk.csv') # or your locale folder
head(PErisk, 2)
```
]

---
class: inverse, center, middle

# select: variable selection

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# select: variable selection

.font130[

* Suppose we want to select only the numeric variables in the `PErisk` dataset.

* We do the following:

```{r}
PErisk_num <- select(PErisk, barb2, gdpw2)
head(PErisk_num, 2)
```
]

---

# select: variable selection

.font130[

The syntax to use the `select` is the following:

```
dat_final <- select(dat_initial, var1, var2, var3, ...)
```
**Your turn**: select the variables `tip` and `totbill` in the `tips` dataset.

]


---

# select: variable selection

.font150[

We can select slices and by characteristics. For example:

```
aux <- select(PErisk, country, gdpw2)

aux <- PErisk %>%
  select(country,gdpw2)

aux <- select(PErisk, -c(courts, prsexp2, prscorr2))
aux <- select(PErisk, courts:prsexp2)
aux <- select(PErisk, starts_with("co"))
```

]

---

# select: variable selection

.font130[

And the methods we can apply in the select are the following:

| Method             | Effect                                          |
|--------------------|-------------------------------------------------|
| v1, v2, v3 (etc)   | Select given variables                          |
| starts_with('xyz') | Select starting with `xyz`                      |
| ends_with('xyz')   | Select ending with `xyz`                        |
| contains('xyz')    | Select variables that have `xyz` in their names |
| vk:vn              | All variables between `vn` and `vk`             |
| -(vk:vn)           | All but `vk` to `vn`                            |

We can even rename variables using `select`.
]

---

class: inverse, center, middle

# rename: change names variables

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# rename: change names variables

.font150[

We can change the names of the variables using `rename`.

Basic syntax:

```
new_dat <- rename(dat, 
                  new_name1 = old_name1,
                  ...)
```

]

---

# rename: change names variables

.font130[

Let's rename the `PErisk` courts to `indep_judiciary`:

```{r}
PErisk_new <- rename(PErisk, indep_judiciary = courts)
head(PErisk_new, 2)
```

**Your turn**: rename the `totbill` to `totalbill` in the dataset `tips`.

]

---


class: inverse, center, middle

# filter: select pieces of the dataset

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# filter: select pieces of the dataset

.font150[

We can select desired rows of the dataset using the function `filter`.

Basic syntax:

```
new_dat <- filter(dat, condition1, condition2, ...)
```

]

---

# filter: select pieces of the dataset

.font130[

Let's filter the `PErisk` to keep all countries without independent judiciary:

```{r}
PErisk_new <- filter(PErisk, 
                     courts == 0)
head(PErisk_new, 2)
```

]

---

# filter: select pieces of the dataset

.font130[

* The filter operators are the following:

```{r}
#| Operator | Meaning                                |
#|----------|----------------------------------------|
#| < and <= | Smaller than and Smaller than or equal |
#| > and >= | Greater than and Treated than or equal |
#| ==       | Equal                                  |
#| !=       | Different                              |
#| !        | Negation                               |
#| |        | Or                                     |
#| &        | And                                    |
```

**Your turn**: filter the `tips` dataset, keeping only the tips above $4.52, by Non-smokers.

]

---


class: inverse, center, middle

# arrange: sort the dataset

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# arrange: sort the dataset

.font150[

We can sort the dataset by levels, according to the needs of our analysis.

Basic syntax:

```
new_dat <- arrange(dat, var1, 
                   desc(var2), 
                   desc(var3), var4, ...)
```
]

---

# arrange: sort the dataset

.font130[

Let's arrange the `PErisk` by the gdpw2, in descending order:

```{r}
PErisk_new <- arrange(PErisk, desc(gdpw2))
head(PErisk_new, 2)
```

**Your turn**: Arrange the `tips` by the size of the tip, in ascending order. See the head and the tail of the arranged data.

]

---

class: inverse, center, middle

# mutate: make transformation in the dataset

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# mutate: make transformation in the dataset

.font150[

We can make transformations in the dataset using the function `mutate`

Basic syntax:

```
new_dat <- mutate(dat, 
                  new_var = do_something_func(old_var),
                  ...)
```

]

---

# mutate: make transformation in the dataset

.font130[

Let's mutate the `PErisk`, to show the gdp in dollars:

```{r}
PErisk_new <- mutate(PErisk,
                     gdppc = exp(gdpw2))
head(PErisk_new, 2)
```

**Your turn**: Let's find the percentage of the tips based on the total bill. Divide the tip by the total bill, and multiply by 100. Sort the data by this variable.

]

---

class: inverse, center, middle

# summarise: summaries of the dataset

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# summarise: summaries of the dataset

.font150[

We can compute summaries of the dataset, performed by a desired variable.

Basic syntax:

```
new_dat <- summarise(dat, 
                     my_summary = do_smt_func(vars),
                     ...)
```

]

---

# summarise: summaries of the dataset

.font130[

Let's compute the mean and standard deviation of the `gdpw2`, and of `barb2`.:

```{r}
PErisk_new <- summarise(PErisk,
                        mean_gdp = mean(gdpw2), std_gdp = sd(gdpw2),
                        mean_barb2 = mean(barb2), std_barb2 = sd(barb2))
head(PErisk_new, 2)
```

**Your turn**: Compute the mean and standard deviation of the tips.

]

---


class: inverse, center, middle

# summarise and group_by: summaries of the dataset by groups

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# summarise and group_by: summaries of the dataset by groups

.font150[

We can compute summaries of the dataset, performed by a desired variable.

Basic syntax:

```
new_dat <- dat %>%
  group_by(var_to_group) %>%
  summarise(my_summary = do_smt_func(vars),
            ...)
```
]

---

# summarise and group_by: summaries of the dataset by groups

.font130[
Let's compute the mean and standard deviation of the `gdpw2`, by expropriation risk:

```{r}
PErisk_new <- PErisk %>% group_by(prsexp2) %>%
  summarise(mean_gdp = mean(gdpw2), std_gdp = sd(gdpw2))
head(PErisk_new, 2)
```
]

---

# summarise and group_by: summaries of the dataset by groups

.font150[
**Your turn**: Compute the mean and standard deviation of the tips, grouped by the smoker and non-smoker.

]

---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>

---

class: inverse, center, middle

# Have a great weekend!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>