---
title: "QTM 151"
subtitle: "Week 5 -- forcats <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: ""
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      self-contained: false
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "https://cloud.r-project.org/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>

# Recap

.font130[

We learned:
 + programming methods
 + `dplyr *_join` methods: joining data
 + `tidyr` methods: reshape datasets

**Great job!!**

Do you have any questions about any of these contents?

Today we are going to talk about **forcats** (package for categorical variables)

]

---

class: inverse, center, middle

# Getting Started

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Getting Started: loading packages

.font130[
```{r}
# Loading tidyverse
library(tidyverse)
```
]

---

# Loading data - GSS

.font130[
```{r, tidy=FALSE}
# Loading the GSS Cat and GSS
gss_cat <- read_csv('https://raw.githubusercontent.com/umbertomig/qtm151/main/datasets/gss_cat.csv')
head(gss_cat, 2)

gss <- read_csv('https://raw.githubusercontent.com/umbertomig/qtm151/main/datasets/gss.csv')
head(gss, 2)
```
]

---

class: inverse, center, middle

# forcats

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# forcats

.font130[
Provide tools to work with categorical data. The methods we will use in here are:

* `fct_reorder()`: reorder levels in a categorical variable
* `fct_relevel()`: move the position of a particular level
* `fct_infreq()`: reorder levels in increasing frequency
* `fct_recode()`: recode factor variable
* `fct_collapse()`: collapse levels in categorical variable
* `fct_lump()`: lump together small groups
* `case_when()`: useful to recode data

]

---

class: inverse, center, middle

# fct_reorder

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# fct_reorder

.font130[

Suppose we want to study the relationship between religion and time watching tv.

This graph here doesn't help much:

```
relig_summary <- gss_cat %>%
  group_by(relig) %>%
  summarise(tvhours=mean(tvhours, na.rm = T)) 

ggplot(relig_summary, aes(x=tvhours, y=relig)) +
  geom_point()
```
]

---

# fct_reorder

.font130[

Reordering the levels of a factor using *fct_reorder()* helps with the interpretation.

*fct_reorder()* takes three arguments:

- f, the factor whose levels you want to modify.

- x, a numeric vector that you want to use to reorder the levels.

- Optionally, fun, a function used if there are multiple values of x for each value of f. The default value is median.

```
ggplot(relig_summary, aes(x=tvhours, y=fct_reorder(relig, tvhours))) +
  geom_point()
```

**Your turn:** plot the marital status by frequency (use the function counts). Then reorder it.

]

---

class: inverse, center, middle

# fct_relevel

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# fct_relevel

.font130[

In the religions x tv hours plot, note that the "Don't know" is the largest category.

It is also not very informative. We can easily send it to the bottom:

```
ggplot(relig_summary, 
       aes(x = tvhours, 
           y = fct_relevel(relig, "Don't know", after=0))) +
    geom_point()
```
]

---

# fct_relevel

.font130[

Can we do ascending, and place the "Don't know" in the bottom?

Yes!

```
ggplot(relig_summary, 
       aes(x=tvhours, 
           y = fct_reorder(relig, tvhours) %>%
             fct_relevel("Don't know", after=0))) + 
  geom_point()
```

**Your turn:** plot the party id by frequency (use the function counts). Then reorder it. Then place "No answer" and "Independent" at the bottom.

]

---

class: inverse, center, middle

# fct_infreq

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# fct_infreq

.font130[

You can use `fct_infreq()` to order levels in increasing frequency:.

* This is the simplest type of reordering because it does not need any extra variables.

You may want to combine with `fct_rev()`.

```
gss_cat %>%
  mutate(marital = fct_infreq(marital)) %>%
  ggplot(aes(marital)) +
    geom_bar()
```
]

---

# fct_infreq + fct_rev (elegant)

.font130[

And that's what happens when we combine both:

```
gss_cat %>%
mutate(marital = marital %>% fct_infreq() %>% fct_rev()) %>%
  ggplot(aes(marital)) + geom_bar()
```

**Your turn:** plot the race by frequency.

]

---

class: inverse, center, middle

# fct_recode

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# fct_recode

.font130[

Recode a categorical variable is always painful, regardless of the statistical software. 

Luckly, the people that wrote forcats made it as easy as it can be by creating the `fct_recode` function.

Check this plot. TV hours by party affiliation:

```
gss_cat %>% 
  drop_na(tvhours) %>% 
  group_by(partyid) %>%
  summarise(meantv=mean(tvhours)) %>%
  ggplot(aes(x=meantv, y=fct_reorder(partyid, meantv)))+
  geom_point()
```
]

---

# fct_recode

.font130[

Note that we can combine some low-information categories together as `others`:

```
gss_cat %>%
  drop_na(tvhours) %>%
  mutate(partyidnew = fct_recode(partyid,
    "Republican, strong"    = "Strong republican",
    "Republican, weak"      = "Not str republican",
    "Independent, near rep" = "Ind,near rep",
    "Independent, near dem" = "Ind,near dem",
    "Democrat, weak"        = "Not str democrat",
    "Democrat, strong"      = "Strong democrat",
    "Other"                 = "No answer",
    "Other"                 = "Don't know",
    "Other"                 = "Other party")) %>%    
  group_by(partyidnew) %>%
  summarise(meantvhours = mean(tvhours)) %>%
  ggplot(aes(x=meantvhours, 
             y=fct_reorder(partyidnew, meantvhours))) +
  geom_point()
```

**Your turn:** plot the health in the gss dataset. Recode to make it interesting.

]

---

class: inverse, center, middle

# fct_collapse

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# fct_collapse

.font130[

`fct_collapse` is good to put several factor levels together.

Look at this code for party id frequency:

```
gss_cat %>%
  mutate(partyidnew = fct_collapse(partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  )) %>%
  count(partyidnew)
```

**Your turn:** collapse the gss `wrkstat` variable to simplify it.

]

---


class: inverse, center, middle

# fct_lump

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# fct_lump

.font130[

`fct_lump` aggregates the small-frequency levels together.

The default is the most frequent + others:

```
gss_cat %>%
  mutate(relignew = fct_lump(relig)) %>%
  count(relignew)
```

]

---


# fct_lump

.font130[

If we want the five most frequent values + others:

```
gss_cat %>%
  mutate(relig = fct_lump(relig, n=5)) %>%
  count(relig)
```

**Your turn:** count the levels in the `gss_cat` `denom` variable. Find a nice way to display it.

]

---


class: inverse, center, middle

# case_when

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# case_when

.font130[

`case_when` is useful for recoding variables. Look at the example below:

```
mtcars %>%
  mutate (
    gear_char =
      case_when(
        gear==3 ~ "three",
        gear==4 ~ "four",
        gear==5 ~ "five"
      )
  )
```

**Your turn:** for `gss_cat`, create a plot to look at how average age varies across income level (rincome)

]

---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>

---

class: inverse, center, middle

# Have a great weekend!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>