---
title: "Homework 1 -- QTM 151"
author: "Your name here"
output: html_document
---

## Instructions:

**Due:** 02/15 at 11:59PM.

**What am I expecting?** An R Markdown and the compiled HTML with the answers.

**Have fun!**

Use `qplot()` function to create plots about `diamonds`. The data set diamonds is built in the package
`ggplot2`.

To get started:

```{r}
library(tidyverse)
data(diamonds)
```

**1**. Make a small random subset of diamonds and name it as `dsmall`.

```{r}
# your answers here
```

**2**. Show the distribution of carat with a histogram, set binwidth to 0.1 and limits for the x-axis from 0
to 3.

```{r}
# your answers here
```

**3**. Show the distribution of carat with a *density plot*, set the color to the variable `color`.

```{r}
# your answers here
```

**4**. Produce a scatter-plot showing the relationship between *carat* and *price*, set point shape to the
variable cut, set point transparency to 0.1, and describe the trend of the scatter-plot.

```{r}
# your answers here
```

**5**. Make another scatter-plot between *carat* and *price*, add jitters to points, set point transparency to
0.1, apply log transformation to the x and y axes variables, and add a smooth line to the plot, then
describe the trend of the scatter-plot.

```{r}
# your answers here
```