---
title: "Homework 3 - dplyr"
author: "Your number here"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(babynames)
library(tidyverse)
library(nycflights13)
```

**About the datasets**:

**`babynames`**: Baby names. Full baby name data provided by the SSA. This includes all names with at least 5 uses. The data frame has five variables: year, sex, name, n and prop (n divided by total number of applicants in that year, which means proportions are of people of that gender with that name born in that year).

**`flights`**: Dataset with flights that departed from NYC (JFL, LGA, or EWR airports). To see the description of the variables type `help(flights)`.

**Q1** (2 points): Calculate the total number of children born each year over time, and then make a line graph to show the trend, and set color to the variable sex.

```{r}
# Your answers here
```

**Q2** (2 points): Identify the top 5 male and female names and show on the barplot, set color for different sex.

```{r}
# Your answers here
```

**Q3** (3 points): For each airport (dest), calculate the total number of flights, the average distance of all flights, and the average arrival delay of all flights, save the results in a dataset (give a name), and remove the airport "HNL" and the airports with less than 20 flights.

```{r}
# Your answers here
```


**Q4** (3 points): Use the dataset you created, provide a scatterplot to show the relationship between travel distance and arrival delay, set size to the total number of flights and the transparency of 0.5, and add a smooth line without showing the SE.

What information can you get from the graph?

```{r}
# Your answers here
```

_Explanation here_