---
title: "Homework 4 - Programming"
author: "Your name here"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

**Instructions**: For the homework, create the required functions, and test them using some examples.

**Q1** (3 points): Write a function that receives a string and return whether the string has more than five letters. Test your function with the two strings provided. (hint: nchar displays the number of characters in a string)

```{r}
# Provided strings
s1 <- 'home' # should be smaller than five characters
s2 <- 'my home' # should be bigger than five characters

# Your answers here

# Testing here
```

**Q2** (3 points): Create a function that receives the weight in pounds and the height in inches and return the BMI. Test with the provided height and weight

```{r}
# Provided data
# height = 72 inches
# weight = 200 pounds
# BMI should be around 27.12

# Your answers here

# Testing here
```

**Q3** (4 points): Create a function that receives a text vector with preferred food, and returns a weekly menu. Test with the vector below. (hint: check out the function `sample`, for the case of sample with replacement.)

```{r}
# Provided foods
foods <- c('pasta', 'pizza', 'apples', 'ice cream')

# Your answers here

# Testing here
```