---
title: "QTM 151"
subtitle: "Week 8 - qplot <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: ""
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "http://cran.cnr.berkeley.edu/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>

# Recap

.font130[

We learned:
 + orogramming methods
 + `dplyr *_join` methods: joining data
 + `tidyr` methods: reshape datasets
 + `forcats` methods: working with categorical variables
 + `lubridate` methods: to work with dates

**Great job!!**

Do you have any questions about any of these contents?

Today we are going to talk about **qplot** (package for quick plots in ggplot2)

Our GitHub page is: https://github.com/umbertomig/qtm151
]

---

class: inverse, center, middle

# qplot

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# qplot

.font130[
* `qplot`, as the names says, stands for quick plots.

* It is great to generate `ggplot` graphs in a hurry.

* And the graphs are indeed `ggplot` objects: you can still add layers to it!

]

---

# qplot - Geoms Available

.font130[
* Which graphs can we generate?
  - **"point"**:scatterplots.
  - **"line"**: line plot.
  - **"histogram"**
  - **"boxplot"**
  - **"density"**
  - **"bar"**: barplot.
  - **"smooth"**: Fits a smooth line.
  - **"dotplot"**: dotplot.
]

---


# qplot - Options for Customization

.font130[
* And there are plenty of quick options to customize the graphs.
  - `data`: Specify the data-frame.
  - `main`: Title.
  - `xlab`, `ylab`: x and y axis labels.
  - `color`: Controls the color of the lines/points.
  - `fill`: Controls the color of areas (e.g. for histograms).
  - `size`: Controls the size of points.
  - `shape`: The shape of points ("circle", "square", "triangle", etc...)
  - `alpha`: Controls the level of transparency of points/lines/fills.
  - `lwd`: Line width.
  - `lty`: Line type ("solid", "dashed", "dotted", etc...).
  - `facets`: Split up the data into multiple plots.
]

---

# Loading tidyverse

.font130[
```{r, tidy=FALSE}
# Load tidyverse
library(tidyverse)
```

]

---

# Loading datasets

.font130[
```{r, tidy=FALSE}
# Loading tips dataset
tips <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm151/main/datasets/tips.csv')
head(tips, 2)

# Loading PErisk dataset
PErisk <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm151/main/datasets/PErisk.csv')
head(PErisk, 2)
```
]

---

class: inverse, center, middle

# Plots for Continuous Variables

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# qplot - Histograms

.font130[

```{r, tidy=FALSE, fig.align='center', fig.height=5, fig.width=5}
qplot(PErisk$gdpw2, geom = "histogram", bins = 10)
```

* **Your turn**: make a histogram of the `tip` variable in the `tips` dataset.

]

---

# qplot - Box-plots

.font130[

```{r, tidy=FALSE, fig.align='center', fig.height=5, fig.width=5}
# Box-plot of log of per capita gdp
qplot(y = PErisk$gdpw2, geom = "boxplot")
```

* **Your turn**: make a boxplot of the `totbill` variable in the `tips` dataset.

]

---

class: inverse, center, middle

# Plot for Discrete Variables

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# qplot - Bar-Plots

.font130[

```{r, tidy=FALSE, fig.align='center', fig.height=5, fig.width=5}
# Bar-plot of courts
qplot(factor(PErisk$courts), geom = "bar")
```

* **Your turn**: make a bar-plot of the `smoker` variable in the `tips` dataset.

]

---

class: inverse, center, middle

# Plot for two Continuous Variables

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# qplot - Scatter-Plot

.font130[

```{r, tidy=FALSE, fig.align='center', fig.height=5, fig.width=5}
qplot(PErisk$gdpw2, PErisk$barb2, geom = "point") + geom_smooth(method = 'lm')
```

* **Your turn**: make a Scatter-Plot for the `totbill` and the `tips` variables in the `tips` dataset.

]

---

class: inverse, center, middle

# Plot for Continuous x Discrete Variables

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# qplot - Multiple Box-Plots

.font130[

```{r, tidy=FALSE, fig.align='center', fig.height=5, fig.width=5}
qplot(x = factor(PErisk$courts), y = PErisk$gdpw2, 
      geom = "boxplot")
```

* **Your turn**: make a box-plot of the `tips` variable by `smoker` in the `tips` dataset.

]

---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 


---

class: inverse, center, middle

# See you next week!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>