---
title: "QTM 151"
subtitle: "Week 2 -- Functions, Loops, and Programming in R <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: ""
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "http://cran.cnr.berkeley.edu/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>

# Before we get started

.font150[

Did you check the GitHub page?

Our GitHub page is: https://github.com/umbertomig/qtm151

Let's check it out?!
]

---

# Today's Agenda

.font130[

`if` statements

`function` creation

`for` loops

]

---

class: inverse, center, middle

# Getting Started

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Getting Started: loading packages

.font130[
```{r}
# Loading tidyverse
library(tidyverse)
```
]

---

# Loading datasets

.font130[
```{r, tidy=FALSE}
# Loading tips dataset
tips <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm150/master/datasets/tips.csv') # or your locale folder
head(tips, 2)

# Loading PErisk dataset
PErisk <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm150/master/datasets/PErisk.csv') # or your locale folder
head(PErisk, 2)

data(USArrests)
```
]

---
class: inverse, center, middle

# if statements

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# if statements

.font130[

**If** statements control the flow of a code.

They create a condition based on some variable and execute a code when the condition is `TRUE`.

Example in pseudo-code:

```
if (number > 10) {
  ...execute code for the bigger than 10 case...
} else {
  ...execute code for the smaller than or equals 10 case...
}
```

]

---

# ifelse

.font130[

In R, we have a function to build if-else statements conveniently.

If-else consists in binary conditions: do something if TRUE, or something else if FALSE.

Syntax: `ifelse(condition, code-if-true, code-if-false)`

For example:

```
n = 4
ifelse(n>2, n^2, n/5)

n = 1
ifelse(n>2, n^2, n/5)

n = 2
ifelse(n>2, n^2, n/5)
```

]

---

# ifelse

.font130[

More examples:

```
## no need to interact with the condition:
n = 2
ifelse(n>2, 'bigger than 2', 'smaller than or equal 2')

## gender: vector operation
gen <- c(0,0,0,1,1,0,1,1,1,0)
ifelse(gen == 1, 'Fem', 'Masc')

## gender: vector operation
gen <- c(0,0,2,1,1,2,1,1,1,0)
ifelse(gen == 1, 'F', ifelse(gen == 0, 'M', 'Non-binary'))
```

**Your turn**: in the tips dataset, take the variable `day`, and create a condition that shows whether the day is weekday or weekend.

]

---

# "if - else if - else" statements

.font130[

Some statements are more extensive than simple ifelse ones.

Example in pseudo-code:

```
if (number > 10) {
  ...execute code for the bigger than 10 case...
} else if (number > 2 & number <= 10) {
  ...execute code for this condition...
} else {
  ...execute code for otherwise...
}
```

]

---

# "if - else if - else" statements

.font130[

Example:

```{r}
n = 20
if(n==1) {
  print('Number equals 1')
} else if (n==2) {
  print('Number equals 2')
} else if (n==3) {
  print('Number equals 3')
} else if (n==4) {
  print('Number equals 4')
} else if (n<0) {
  print('Number smaller than zero')
} else {
  print('Invalid number!')
}
```

]

---

class: inverse, center, middle

# functions

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# functions

.font130[

Sometimes we need to create our functions.

Creating functions is simple. The syntax is straightforward:

```
name_function <- function(par1, par2, par3, ...) {
  ...code execute...
  return(...return smt...)
}
```

Example in pseudo-code:

```
testn <- function(n) {
  if (number > 10) {
    ...execute code for the bigger than 10 case...
  } else {
    ...execute code for the smaller than or equals 10 case...
  }
  return(...smt...)
}
```

]

---


# functions

.font130[

Example: BMI index function

```{r}
bmi <- function(w, h) {
  bmi = w / (h^2)
  return(bmi)
}
# BMI person 70 kg and 1.75 m
bmi(70, 1.75)
```

**Your turn** create a function that receives weight in pounds and heights in inches and returns the BMI.
]

---

class: inverse, center, middle

# loops

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# loops

.font130[

There are two ways of creating loops: `while` and `for`.

While loops are useful to run a piece of code until a condition is satisfied. 

For example:

```
i = 1
while(i < 50) {
  print(i)
  print('i is smaller than 50!')
  print(i^10)
  i = i+1
}
```

]


---

# loops

.font130[

Example: infinite loop:

```
while(TRUE) {
  x = readline("Type something or break: ")
  if(x == 'break'){
    break
  }
  cat("You typed: ", x)
}
```

]

---

# loops

.font130[

The `for` loop is a bit different: it receives a counter and runs until it is exhausted.

Example:

```
for (i in 1:10) {
  print(i)
  print(i^5)
}
```

]

---

# loops

.font130[

Example: let's run over the columns of a dataset and display the class of each variable:

```
for (i in names(tips)) {
  print(class(tips[,i]))
}
```

**Your turn**: Create a function that displays the summary of the numeric or integer variables and a table of the other variables.

]

---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>

---

class: inverse, center, middle

# Have a great weekend!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>