---
title: "QTM 151"
subtitle: "Week 4 -- tidyr <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: ""
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      self-contained: false
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "https://cloud.r-project.org/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>

# Recap

.font130[

We learned:
 + Programming: `for`, `while`, conditional statements
 + `dplyr *_join` methods: joining data

Do you have any questions about any of these contents?

Our GitHub page is: https://github.com/umbertomig/qtm151
]

---

class: inverse, center, middle

# Getting Started

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Getting Started: loading packages

.font130[
```{r}
# Loading tidyverse
library(tidyverse)
```
]

---

# Loading data

.font130[
```{r, tidy=FALSE}
# population table
table1<-tibble(
  `country`=c("Afghanistan","Brazil","China"),
  `1999`=c(19987071, 172006362, 1272915272),
  `2000`=c(20595360, 174504898, 1280428583)
)
table1

# case table
table2<-tibble(
  `country` = c("Afghanistan","Brazil","China"),
  `1999` = c(745, 37737, 212258),
  `2000` = c(2666, 80488, 213766)
)
table2
```
]

---

# Loading data

.font130[
```{r, tidy=FALSE}
table3<- tibble(
  `country` = c("Afghanistan", "Afghanistan", "Afghanistan", "Afghanistan", "Brazil", "Brazil", "Brazil", "Brazil","China", "China","China", "China"),
   `year` = c(1999,1999,2000,2000,1999,1999,2000,2000,1999,1999,2000,2000),
   `type` = c("case", "population","case", "population","case", "population","case", "population","case", "population","case", "population"),
   `count` = c(745, 19987071, 2666, 20595360, 37737, 172006362, 80488, 174504898, 212258, 1272915272, 213766, 1280428583)
)
table3
```
]

---

# Loading data

.font130[
```{r, tidy=FALSE}
table4<-tibble(
    `country` = c("Afghanistan", "Afghanistan", "Brazil", "Brazil","China", "China"),
     `year` = c(1999,2000,1999,2000,1999,2000),
    `rate`= c("745/19987071", "2666/20595360", "37737/172006362", "80488/174504898", "212258/1272915272", "213766/1280428583")
)
table4
```
]

---

# Loading data

.font130[
```{r, tidy=FALSE}
# population table
tbl1<-tibble(
  `county`=c("DeKalb","Fulton","Cobb"),
  `2010`=c(691961, 920581, 690688),
  `2011`=c(693961, 921581, 691688)
)
tbl1

# veterans table
tbl2<-tibble(
  `county` = c("DeKalb","Fulton","Cobb"),
  `2010` = c(36189, 42448, 41345),
  `2011` = c(36389, 42648, 41545)
)
tbl2
```
]

---

# Loading data

.font130[
```{r, tidy=FALSE}
tbl3<- tibble(
  `country` = c("DeKalb", "DeKalb", "DeKalb", "DeKalb", "Fulton", "Fulton", "Fulton", "Fulton","Cobb", "Cobb","Cobb", "Cobb"),
   `year` = c(2010,2010,2011,2011,2010,2010,2011,2011,2010,2010,2011,2011),
   `type` = c("veterans", "population","veterans", "population","veterans", "population","veterans", "population","veterans", "population","veterans", "population"),
   `count` = c(36189, 691961, 36389, 693961, 42448, 920581, 42648, 921581, 41345, 690688, 41545, 691688)
)
tbl3
```
]

---


# Loading data

.font130[
```{r, tidy=FALSE}
tbl4<- tibble(
  `country` = c("DeKalb", "DeKalb", "Fulton", "Fulton", "Cobb", "Cobb"),
   `year` = c(2010,2011,2010,2011,2010,2011),
   `prop` = c("36189/691961", "36389/693961", "42448/920581", "42648/921581", "41345/690688", "41545/691688")
)
tbl4

stocks1 <-tibble(
  year = rep(c(2015,2016), each=4),
  qtr = rep(c(1,2,3,4), 2),
  return = c(1.88, 0.59, 0.35, NA, NA, 0.92,0.17, 2.66)
) %>% na.omit()
```
]

---

class: inverse, center, middle

# tidyr

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# tidyr

.font130[
The *tidyr* package helps tidy up messy datasets. There are three interrelated rules which make a dataset tidy:

1. Each variable must have its own column

2. Each observation must have its own row

3. Each value must have its own cell

There are a few key functions in the *tidyr* package, *gather()*, *spread()*, *separate()*, *unite()*, *complete()*, *fill()*.
]

---

class: inverse, center, middle

# gather

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# gather

.font130[

To tidy  a dataset, we need to *gather* multiple columns, and gathers them into key-value pairs: it makes "wide" data longer.

Syntax:
```
gather(dataset, set_of_columns, 
       key="name of variable whose values form the column names", 
       value="name of variable whose values are spread over the cells")
```
]

---

# gather

.font130[
Example:

```{r}
gather(table1, `1999`,`2000`, key="year", value="population")
```

]

---

# gather

.font130[
Example:

```{r}
gather(table2, "1999":"2000", key=year, value = cases)
```

**Your Turn**: Do the same with the `tbl1` and `tbl2` datasets. Save the results and join the datasets.

]

---

class: inverse, center, middle

# spread

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# spread

.font130[

*spread()* is the opposite of *gather()*.

*gather()* makes wide tables narrower and longer, *spread()* makes long tables shorter and wider.

Syntax:
```
spread(dataset, 
       key="the column that contains variable name", 
       value="the column that contains values forms multiple variables")
```
]

---

# spread

.font130[
Example:

```{r}
spread(table3, key= "type", value= "count")
```

**Your Turn**: Check out `tbl3`. Then, spread it by `type`.

]

---

class: inverse, center, middle

# separate and unite

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# separate and unite

.font130[

*separate()* pulls apart one column into multiple columns, *unite()* is the inverse of *separate()*.

Check `table4`. Note that the rate variable has two variables inside it: cases and population. To separate them:

Syntax for *separate()*:
```
separate(table, variable_separate, into=c('v1','v2'), sep="/")
```
]

---

# separate and unite

.font130[
Example:

```{r}
table5 <- separate(table4, rate, into=c('case','population'), sep="/")
table5
```

**Your Turn**: Do the same separate for `tbl4`.

]

---

# separate and unite

.font130[
We can also separate by position:

```{r}
separate(table3, year, into = c("century", "year"), sep=2)
```

**Your Turn**: Do the same to separate `tbl3` year by position.

]

---

# separate and unite

.font130[

Syntax for *unite()*:
```
unite(table, name_col, col1, col2, ..., sep="/")
```
]

---

# separate and unite

.font130[
Example:

```{r}
unite(table5, col = rate, case, population, sep="/")
```

]

---


class: inverse, center, middle

# complete

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# complete

.font130[

*complete()* is useful to fill up the columns with missing data, based on a given pattern.

Suppose we have the following dataset:

```{r}
df <- tibble(
  group = c(1:2, 1), item_id = c(1:2, 2),
  item_name = c("a", "b", "b"),
  value1 = 1:3, value2 = 4:6
)
df
```
]

---

# complete

.font130[
We can complete by group, item_id and item_name:

```{r}
df %>% complete(group, nesting(item_id, item_name))
```

**Your Turn**: Do the same complete for `stocks1`. What happened?

]

---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>

---

class: inverse, center, middle

# Have a great weekend!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>