---
title: "QTM 151"
subtitle: "Week 3 -- plotly <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: ""
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "https://cloud.r-project.org/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>

# Recap

.font130[

* Last week we learned `qplot`. It is a great function simple dataviz.

* Do you have any questions about `qplot`?

* I also learned that some of you have difficulty to download data from GitHub. I posted a video to help with this.

* The quiz is already posted, and I'll post the 

* Our GitHub page is: https://github.com/umbertomig/qtm151
]

---

# Today's Agenda

.font150[

* Plotly

* Creating `plotly` from ggplot graphs

* Using the `plotly` interactive device 

]

---

class: inverse, center, middle

# plotly

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# plotly

.font150[
* `plotly` is an R package for creating interactive web-based graphs.

* It is based on `plotly.js`, a javascript to plot interactive graphs in R and Python.

* They are great for presentations and for displaying data for a non-stats crowd.

]

---

# plotly

.font150[

* There are two functions that you can use to create a plotly:
  + The `plot_ly()` function, that transforms data into a `plotly` object.
  + The `ggplotly()` function, that transforms a ggplot object into a `plotly` object.
  
* Regardless of how a `plotly` object is created, printing it results in an interactive web-based visualization enabled by default. 

]

---

# Getting Started: loading packages

.font130[
```{r}
# Loading tidyverse
library(tidyverse)

# Loading plotly
library(plotly)
```
]

---

# Loading datasets

.font130[
```{r, tidy=FALSE}
# Loading tips dataset
tips <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm151/main/datasets/tips.csv')
head(tips, 2)

# Loading PErisk dataset
PErisk <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm151/main/datasets/PErisk.csv')
head(PErisk, 2)

# Load gapminder
library(gapminder)
head(gapminder, 2)
```
]

---

# Loading datasets

.font130[
```{r, tidy=FALSE}
# Load gapminder
library(gapminder)
head(gapminder, 2)
```
]

---

# ggplot

```{r, fig.align='center', fig.height=5, fig.width=5}
p <- ggplot(PErisk, aes(x=barb2, y=gdpw2)) + 
  geom_point() +
  geom_smooth(method="lm", se=F)

p
```

---

# plotly

```{r, fig.align='center', message = FALSE}
ggplotly(p)
```

---

# ggplotly

.font150[
* **Your turn:** Create a `ggplotly` with `totbill` and `tip`.

* We can do even better: create a plotly with colors by factor.

]

---

# ggplot

```{r, fig.align='center', fig.height=5, fig.width=5}
p <- ggplot(PErisk, aes(x=barb2, y=gdpw2)) + 
  geom_point(aes(col = factor(courts))) +
  geom_smooth(aes(col = factor(courts)), method="lm", se=F)
p
```

---

# plotly

```{r, fig.align='center', message = FALSE, echo = FALSE, warning=FALSE}
ggplotly(p)
```

---

# Intro to Animations in ggplot2

.font130[

* Animations can be created by using the *frame* argument aesthetic in ggplotly().

* By default, animations populate a play button and slider component for controlling the state of the animation (to pause an animation, click on a relevant location on the slider bar).

* [One example](https://www.youtube.com/watch?v=hVimVzgtD6w)

]

---

# plotly

```{r, fig.align='center', message = FALSE, echo = FALSE}
p <- ggplot(gapminder, aes(x=gdpPercap, y=lifeExp, 
                           color = continent)) +
  geom_point(aes(size = pop, 
                 frame = year, ids=country)) +
  scale_x_log10()

ggplotly(p)
```
---

# ggplotly

.font150[
* **Your turn:** Create a plotly with `totbill` and `tip`, varying by day and time.

* And you can keep adding animation options:
  + transition: the duration of the smooth transition between frame (in milliseconds)
  + easing: the type of transition easing
  + mode: describes how a new animate call interacts with currently-running animations.
]

---


# ggplotly: more animation options

```{r, echo = FALSE, warning=FALSE}
ggplotly(p) %>%
  animation_opts(transition = 500, 
                 easing = "linear", 
                 mode = "immediate")
```

---

# ggplotly: more animation options

```{r, echo = FALSE, warning=FALSE}
ggplotly(p) %>% 
  animation_opts(
    1000, easing = "elastic", redraw = FALSE
  ) %>% 
  animation_button(
    x = 1, xanchor = "right", y = 0, yanchor = "bottom"
  ) %>%
  animation_slider(
    currentvalue = list(prefix = "YEAR ", font = list(color="red"))
  )
```

---

# ggplotly

.font150[
* **Your turn:** Make some animation options in your previous plot: `totbill` x `tip`, varying by day and time.
]

---

# plot_ly

.font150[
* The *plot_ly()* function provides a direct interface to plotly.js, so anything in the [figure reference](https://plot.ly/r/reference/) can be specified via *plot_ly()*.

* A plotly visualization is composed of one (or more) trace(s), and every trace has a *type* (the default trace type is "scatter") can be used to draw a large amount of geometries along with the *add_XX()* functions.

* The *plot_ly()* function has a number of arguments that make it easier to scale data values to visual aesthetics (e.g., *color/colors, symbol/symbols, linetype/linetypes, size/sizes*). 

* Template:
```
plot_ly(data, x, y, symbol, size,  type, mode, color)
```

]

---


# plot_ly

```
p<-plot_ly(tips, x=~tip, type="histogram")

p1<-plot_ly(tips, x=~tip) %>% add_histogram(name="tip")
p2<-plot_ly(tips, x=~totbill) %>% add_histogram(name="totalbill")

subplot(p1,p2) %>% hide_legend()
```

---


# plot_ly

```{r, echo = FALSE, warning=FALSE}
p<-plot_ly(PErisk, x=~gdpw2, type="histogram")
p
```

---

# plot_ly

```{r, echo = FALSE, warning=FALSE}
plot_ly(PErisk, x=~gdpw2) %>% 
  add_histogram(name="GDP Per Capita")
```

---


# plot_ly

```{r, echo = FALSE, warning=FALSE}
plot_ly(PErisk, x=~barb2) %>%
  add_histogram(name="Informal Markers Gains")
```

---


# plot_ly

```{r, echo = FALSE, warning=FALSE}
p1<-plot_ly(PErisk, x=~gdpw2) %>% add_histogram(name="GDP Per Capita")
p2<-plot_ly(PErisk, x=~barb2) %>% add_histogram(name="Informal Markers Gains")
subplot(p1,p2) %>% hide_legend()
```

---

# plot_ly

.font150[
* **Your Turn**: do the same plots for `tip` and `totbill`.

* We can also do boxplots.
]

---

# plot_ly: boxplot

```{r, echo = FALSE, warning=FALSE}
plot_ly(PErisk, x = ~courts, y=~barb2, type = "box")
```
---
# plot_ly: boxplot

```{r, echo = FALSE, warning=FALSE}
plot_ly(PErisk, x = ~prsexp2, y=~gdpw2) %>% add_boxplot()
```

---

# plot_ly: boxplot

```{r, echo = FALSE, warning=FALSE}
plot_ly(PErisk, x=~prsexp2, y=~gdpw2, type="box",
        boxpoints="all", pointpos=1,
        marker=list(color="green"))  %>%
  layout(title = "Boxplot GDP by Expropriation Risk")
```

---


# plot_ly: scatterplots

```{r, echo = FALSE, warning=FALSE}
plot_ly(PErisk, x=~barb2, y=~gdpw2, 
        type="scatter", mode="markers") 
```

---

# plot_ly: scatterplots

```{r, echo = FALSE, warning=FALSE}
plot_ly(PErisk, x=~barb2, y=~gdpw2, 
        type="scatter", mode="markers") %>%
  add_markers(symbol= ~prsexp2) 
```

---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>

---

class: inverse, center, middle

# Have a great weekend!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>