---
title: "QTM 151"
subtitle: "Lab 08 -- Shiny web applications <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: "Summer"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      self-contained: false
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "https://cloud.r-project.org/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>

# Recap

.font130[

We learned:
 + `qplot`: quick way to make ggplot graphs.
 + `ggplotly` and `plot_ly`: create nice plotly graphs.
 + `dplyr *_join` methods: joining data
 + `forcats` methods: working with categorical variables
 + `lubridate` methods: processing dates and times
 + `tidyr` methods: reshape datasets
 + `maps` and `ggmap` for creating maps.

**Great job!!**

Do you have any questions about any of these contents?

Today we are going to talk about Shiny Web Apps in R.

]

---

# This Class

.font130[

We have a quiz for this class, due in a few days. Please let me know if you have any questions.

How is the final project going? Do you need my help?

Our GitHub page is: https://github.com/umbertomig/qtm151
]

---

class: inverse, center, middle

# Shiny Webapps

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Getting Started: loading packages

.font130[
```{r}
# Loading tidyverse
library(tidyverse)
library(shiny)
```
]

---


# Getting Started: loading data

.font130[
We will use the shiny presentation by Garrett Grolemund, one of the R Studio main Data Scientists
]

---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>

---

class: inverse, center, middle

# Thank you so much for the class! I hope to see you soon in my QTM 385 class!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>