---
title: "Midterm"
author: "Your name here"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Question 1 (1 pt) - Loading Datasets and Packages

On the GitHub page, I placed five datasets:

1. `corona.csv`
2. `partial1.csv`
3. `partial2.csv`
4. `vaccinesUS.csv`
5. `vaccinesWorld.csv`

Also load the following packages:

1. `tidyverse`
2. `plotly`

```{r}
# Your answers here
```

## Question 2 (1 pt) - qplot

Experts claim that the COVID death rate varies with the number of people with underlying diseases, such as diabetes. In the dataset `corona` you have the variables `cvd_death_rate` and `diabetes_prevalence`. Plot the diabetes prevalence against the covid death rate, to see the relationship. Add a smooth linear trend line. Use the function `qplot`. Would you say that this data corroborates this hypothesis?

```{r}
# Your answers here
```

_No relations_

## Question 3 (2 pt) - ggplotly

Using the `corona` dataset, make the previous plot (plot from question 2), but now displaying the values for the observations. Use `ggplotly`, and reuse the previous code.

```{r}
# Your answers here
```

## Question 4 (1 pt) - plot_ly

The dataset `vaccinesUS.csv` has the daily vaccinations the US. Make a plot that displays the increase of vaccination each day. Use the function `plot_ly`. (hint: look at the first plot in here: https://plotly.com/r/time-series/)

```{r}
# Your answers here
```

## Question 5 (2 pt) - dplyr

For the dataset `corona`, compute the average of the following variables: `cvd_death_rate`, `aged_70_older`, `diabetes_prevalence`, `female_smokers`, `male_smokers` by continent. Arrange by the mean of the covid death rate, in descending order. Use group_by and summarize. Use pipe (`%>%`).

```{r}
# Your answers here
```

## Question 6 (1 pt) - dplyr join

Join the datasets `partial1.csv` and `partial2.csv`. You should use `full_join`. Hint: to join the datasets you will need to find the key variable and match it. When done, plot the head of the end dataset.

```{r}
# Your answers here
```

## Question 7 (2 pt) - tidyr

The dataset `vaccinesWorld.csv` head is like this, for the variables `location`, `date`, and `total_vaccinations`:

```
# A tibble: 6 x 3
  location    date       total_vaccinations
  <chr>       <date>                  <dbl>
1 Afghanistan 2021-02-22                  0
2 Afghanistan 2021-02-23                 NA
3 Afghanistan 2021-02-24                 NA
4 Afghanistan 2021-02-25                 NA
5 Afghanistan 2021-02-26                 NA
6 Afghanistan 2021-02-27                 NA
```

Use tidyr to transform it into this:

```
# A tibble: 182 x 117
   location            `2020-12-13` `2020-12-14` `2020-12-15`
   <chr>                      <dbl>        <dbl>        <dbl>
 1 Afghanistan                   NA           NA           NA
 2 Africa                        NA           NA           NA
 3 Albania                       NA           NA           NA
 4 Algeria                       NA           NA           NA
 5 Andorra                       NA           NA           NA
```

```{r}
# Your answers here
```